---
title: "Lab2"
author: "Ziwen Zhang"
date: "2022-11-10"
output:
  word_document:
    fig_height: 5.5
    fig_width: 10.0
  pdf_document: 
    latex_engine: xelatex
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```
# Question 1
```{r}
# read the file
arthritis <- read.csv("arthritis.csv")

# reordered the "Improved" column
Improvedord <- ordered(arthritis$Improved, levels=c("None", "Some", "Marked"))

# create a new data frame with the reordered Improved column
new <- data.frame(arthritis, Improvedord)

# make a 3-way table
mytable <- xtabs(~ Treatment+Sex+Improvedord, data=arthritis)
ftable(mytable)

```

# Question 2
```{r}
# create a matrix with aspirin values
aspirin <- matrix(c(189, 104, 10845, 10933), nrow = 2)
aspirin

# write a function to calculate odds ratio, relative risk ratio and difference of proportions for a 2 x 2 table.
calculate <- function(data){
  odds1 <- data[1,1] / data[1,2]
  odds2 <- data[2,1] / data[2,2]
  odds_ratio <- odds1 / odds2
  
  relative_risk1 <- data[1,1]/(data[1,1] + data[1,2])
  relative_risk2 <- data[2,1]/(data[2,1] + data[2,2])
  relative_risk_ratio <- relative_risk1 / relative_risk2
  
  diffProp <- data[1,1]/(data[1,1] + data[1,2]) - data[2,1]/(data[2,1] + data[2,2])
  
  paste("odds ratio: ", round(odds_ratio,3), "relative risk ratio: ", round(relative_risk_ratio,3), "difference of proportions: ", round(diffProp,3))
}

# apply the function on aspirin data
calculate(aspirin)
```

# Question 3
```{r}
# set gama values
gama <- c(0.1, 0.01, 0.001, 0.0001)

# calculate estimated probability
pEst <- 0.85*gama / (0.85*gama + 0.01*(1-gama))

# create a data frame to combine gama values and the results
result <- data.frame(gama, pEst)

# plot the result
plot(result$gama, result$pEst, 
     main = "The estimated probability that subject truely has the disease (given positive test)", 
     xlab = "γ", 
     ylab = "Estimated probability")
```

As the probability of a subject has the disease (γ values) increases, the estimated probability that subject truly has the disease (given positive test) increases as well. The trend is more significant when γ values are small.

# Question 4
Consider an emergency room that, on the average, accepts 5 patients per 3 hours.
• What is the probability that there will be no patients next hour?
• At least two patients next hour?
```{r}
# set Y as the number of patients per hour, then Y has Poisson distribution with λ = 5/3

prob1 <- dpois(0, 5/3, log=FALSE)
paste("The probability that there will be no patients next hour is ", round(prob1, 4))

prob2 <- ppois(2-1, 5/3, lower.tail = FALSE, log.p = FALSE)
paste("The probability that at least two patients next hour is ", round(prob2, 4))
```

