---
title: "Lab 5"
author: "Ziwen Zhang"
date: "2022-12-08"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1
Consider data tongue from the package KMsurv. Do a simple graphical check of the proportional hazards assumptions and state your conclusions.
```{r}
library(KMsurv)
library(survival)
library(survminer)
data(tongue)

tongue <- tongue %>%
  mutate(type = ifelse(type == 1, "Aneuploid Tumor", "Diploid Tumor"))

tongue_fit <- survfit(Surv(time, delta) ~ type, data = tongue)

ggsurvplot(tongue_fit, fun = "cloglog", data = tongue) + 
  labs(title = "Log-log curves by tumor type",
       y = "log-log survival", 
       x = "Time to death or on-study time (in weeks)")
```

The proportional hazards assumption is violated since the log-log survival curves are not parallel.


# Question 2
For this problem we use data kidtran (death times of 863 kidney transplantpatients) from KMsurv package. Patients can be classified by race and sex into four groups.

• Test the hypothesis that there is no difference in survival between the four groups.
• for each gender group, conduct individual tests of the hypothesis of no racial differences in survival rates.

```{r}
data("kidtran")

log_rank <- survdiff(Surv(time, delta) ~ race + gender, data = kidtran)
log_rank
```
H0: There is no difference in survival between the four groups.
H1: There is at least one difference in survival among the four groups.

Since the p-value is 0.2 > 0.05, there is no evidence to reject the null hypothesis and thus there is no difference in survival among the four groups.

```{r}
library(tidyverse)

# for male (gender=1)
male <- kidtran %>%
  filter(gender == 1)

log_rank_male <- survdiff(Surv(time, delta) ~ race, data = male)
log_rank_male

# for female (gender=2)
female <- kidtran %>%
  filter(gender == 2)

log_rank_female <- survdiff(Surv(time, delta) ~ race, data = female)
log_rank_female
```

For male group:

H0: There is no difference in survival between the two race groups.
H1: There is a difference in survival between the two race groups.

Since the p-value is 0.8 > 0.05, there is no evidence to reject the null hypothesis and thus there no difference in survival between the two race groups.

For female group:

H0: There is no difference in survival between the two race groups.
H1: There is a difference in survival between the two race groups.

Since the p-value is 0.03 < 0.05, there is an evidence to reject the null hypothesis and thus there is a difference in survival between the two race groups.

# Question 3
For the female subset from the previous problem (kidtran) estimate the difference in RMST for τ = 2500 using package survRM2.
```{r}
library(survRM2)

female <- female %>%
  mutate(race = ifelse(race == 1, 0, 1))

RMSTdiff <- rmst2(time = female$time, status = female$delta, arm = female$race, tau = 2500, covariates = NULL, alpha = 0.05)
RMSTdiff
```

The difference in RMST is 216.054

# Question 4
The cumulative hazard function for the Weibull distribution is given by
$$
H(X) = λx^α
$$
Derive a simple graphical diagnostic to check whether the Weibull distribution is adequate for data.

Please note that this question is not about specific data set. Your solution should be applicable for any data when you want to use Weibull distribution

Solution:
Take the logarithm to the left and right of the equation:
$$
log(H(x)) = log(λx^α) = log(λ) + log(x^α) = log(λ) + a*log(x)
$$
Therefore, we can calculate $$H(x)$$ and plot $$log(H(x))$$ versus $$log(x)$$. If the data can fit in this linear model, the Weibull distribution is adequate for data.